---
title: "p8105_hw2_kh3074"
output: github_document
time: 2021.09.30
author: KaiYu He(kh3074)
---


---
```{r}
library(tidyverse)
library(readxl)
```

## \textcolor{blue}{Problem1}  

#### Read and clean the Mr. Trash Wheel sheet:  

```{r}
trash_weel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx") %>%
  janitor::clean_names()%>% #change to reasonable variable names
  select(dumpster:homes_powered)%>% #omit the columns with notes
  filter(dumpster != 'NA')%>% #omit rows that do not include dumpster-specific data
  mutate(sports_balls = round(sports_balls))#round the number of sports balls to the nearest integer
trash_weel_df
summary(trash_weel_df)
```

#### precipitation data for 2018 and 2019

```{r}
precipitation_2018 = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2018 Precipitation",skip = 1)%>%
  janitor::clean_names()%>%
  mutate('year' = 2018)%>%
  drop_na()

  
precipitation_2019 = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2019 Precipitation", skip = 1)%>%
  janitor::clean_names()%>%
  mutate('year' = 2019)%>%
  drop_na()
  
precipitation_2018_2019 = 
  full_join(precipitation_2018,precipitation_2019)%>% 
  mutate(month = month.name[month])
precipitation_2018_2019
summary(precipitation_2018_2019)
```

#### Conclusion
- **Trash weel sheet**  
  Form `r nrow(trash_weel_df)` rows and `r ncol(trash_weel_df)` columns of data,  
  we can learn that Mr trash weel collect `r mean(pull(trash_weel_df,'cigarette_butts'))` cigarette butts per dumpster which is the most among all the trashes,  
  collect `r mean(pull(trash_weel_df,'weight_tons'))` tons of trash per dumpster.   
  And the median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trash_weel_df,year == 2017),sports_balls))`
  

- **Precipitation sheet**  
  From `r nrow(precipitation_2018_2019)` rows and `r ncol(precipitation_2018_2019)` columns of data,  
  we can find out the mean precipitation per month is `r mean(pull(precipitation_2018_2019,'total'))`,  
  total precipitation in 2018 is `r sum(pull(filter(precipitation_2018_2019,year == 2018),total))`.
  
  
