---
title: "p8105_hw2_kh3074"
output: github_document
time: 2021.09.30
author: KaiYu He(kh3074)
---


---
```{r}
library(tidyverse)
library(readxl)
```

---

# Problem1

#### Read and clean the Mr. Trash Wheel sheet:  

```{r}
trash_weel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx") %>%
  janitor::clean_names()%>% #change to reasonable variable names
  select(dumpster:homes_powered)%>% #omit the columns with notes
  filter(dumpster != 'NA')%>% #omit rows that do not include dumpster-specific data
  mutate(sports_balls = round(sports_balls))#round the number of sports balls to the nearest integer
trash_weel_df
summary(trash_weel_df)

```

#### precipitation data for 2018 and 2019

```{r}
precipitation_2018 = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2018 Precipitation",skip = 1)%>%
  janitor::clean_names()%>%
  mutate('year' = 2018)%>%
  drop_na()

  
precipitation_2019 = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2019 Precipitation", skip = 1)%>%
  janitor::clean_names()%>%
  mutate('year' = 2019)%>%
  drop_na()
  
precipitation_2018_2019 = 
  full_join(precipitation_2018,precipitation_2019)%>% 
  mutate(month = month.name[month])
precipitation_2018_2019
summary(precipitation_2018_2019)
```

#### Conclusion
- **Trash weel sheet**  
  Form **`r nrow(trash_weel_df)`** rows and **`r ncol(trash_weel_df)`** columns of data,  
  we can learn that Mr trash weel collect **`r mean(pull(trash_weel_df,'cigarette_butts'))`** cigarette butts per dumpster which is the most among all the trashes,  
  collect **`r mean(pull(trash_weel_df,'weight_tons'))`** tons of trash per dumpster.   
  And the median number of sports balls in a dumpster in 2017 is **`r median(pull(filter(trash_weel_df,year == 2017),sports_balls))`**
  

- **Precipitation sheet**  
  From **`r nrow(precipitation_2018_2019)`** rows and **`r ncol(precipitation_2018_2019)`** columns of data,  
  we can find out the mean precipitation per month is **`r mean(pull(precipitation_2018_2019,'total'))`**,  
  total precipitation in 2018 is **`r sum(pull(filter(precipitation_2018_2019,year == 2018),total))`**.
  
# Problem 2

#### Clean DataSet of pols-month.csv
```{r}
pols_month =
  read_csv("data/fivethirtyeight_datasets/pols-month.csv")%>%
  separate(mon, into = c("year", "month", "day"), sep="-") %>%
  mutate(month = as.double(month))%>%  #change the variable type to double
  mutate(month = month.name[month])
pols_month

pols_month_tidy = 
  pols_month%>%
  mutate(president = prez_dem+prez_gop)%>%
  select(-'day',-'prez_gop',-'prez_dem')
pols_month_tidy
```
#### Clean data snp.csv
```{r}
snp_df = 
  read.csv("data/fivethirtyeight_datasets/snp.csv")%>%
  separate(date,into = c('month','day','year'),sep ='/')%>%
  relocate(year)%>%
  select(-day)%>%
  mutate(month = as.numeric(month))%>% #Change the month name format 
  mutate(month = month.name[month])

unemp_df = 
  read.csv("data/fivethirtyeight_datasets/unemployment.csv")%>%
  pivot_longer(
    Jan:Dec,
    names_to = 'month',
    values_to = 'unemployment'
  )%>%
  mutate(month = match(month,month.abb))%>% #Change the month name format 
  mutate(month = month.name[month])%>%
  mutate(Year = as.character(Year))%>%  #make sure the datatype are the same
  janitor::clean_names()

#Join the datasets by merging snp into pols
pols_snp_df = left_join(pols_month_tidy,snp_df,by = c('year','month'))
pols_snp_df
knitr::kable(tail(pols_snp_df)) # show some of the results


#Merging unemployment into the result.
pols_snp_unemployment_df = left_join(pols_snp_df,unemp_df,by = c('year','month'))
pols_snp_unemployment_df
knitr::kable(tail(pols_snp_unemployment_df))#show some of the results
```

#### Data interpretation 
- pols_month_df is a **`r nrow(pols_month_tidy)`** rows and **`r ncol(pols_month_tidy)`** columns data which include numbers of government staffs grouped by gov sen and rep.
- snp_df is a **`r nrow(snp_df)`** rows and **`r ncol(snp_df)`** columns data record the closing values of the S&P stock index on the associated date.
- unemployment_df is a **`r nrow(unemp_df)`** rows and **`r ncol(unemp_df)`** columns data include numbers of percentage of unemployment on the associated date.
- Combine all the data togather we can analyze the relationship between the proportion of government staffs and S&P stock index and unemployment.
